A game theoretic model of the behavioural gaming that takes place at the EMS - ED interface\tnoteref{t1}

### 2022-04-05 sections/introduction/main.tex
Add a final TODO to be fixed
Changes after second supervisory review (#45)
* Add changes to response to reviewers

* Fix square in general markov model image

* Add another keyword

* Apply changes to intro and lit review

* Apply changes to queueing mode section

* Apply changes to game theoretic mode section

* Apply changes to conclusion

* Minor changes to appendix

* Compile PDF

There are numerous news articles that focus on the complexity that arises when
ambulances stay blocked outside of the hospital for a long amount of 
time~\cite{dailyrecords, staffordshirelive}.
Some news reports comment on the long idle time of ambulances when
they are not in use~\cite{herefordtimes} and there are several reports of 
examples where this became an issue for new patients~\cite{southwalesargus}.
A particular article looked into this problem from the point of view of an 
ambulance paramedic:

\begin{quotation}
    % <!--alex ignore he-she her-him-->
    ``He knows if a patient arrives more urgently in need of help, his will be 
    pushed back''~\cite{bbcwales}

    ``We used to bring poorly patients in, and we were out on the road again in
    15 minutes.''~\cite{bbcwales}
\end{quotation}

% ED + EMS application
This paper aims to describe the EMS-ED interface using a game theoretic model
informed by an underlying queueing model.
The model describes the situation where an ambulance service would
have to distribute its patients between two EDs.
The two EDs can be thought of as two queueing systems and the EMS as a 
distributor that distributes patients to them, aiming to minimise some
performance measure.
The patients that are distributed by the EMS arrive at the hospital via an 
ambulance and are then either offloaded at the ED or stay blocked outside in 
the ambulance.
Whether or not the ambulance and its patient stay blocked is determined by 
the threshold that the given ED chooses to play.
High threshold indicates that the ED accepts ambulance patients even if it is 
relatively full, while low threshold means that the ED blocks ambulances more 
frequently.
In the United Kingdom the National Heath Services (NHS) sets some regulations 
on ED performance.
One of these regulations is that 95\% of patients that arrive at the ED should 
be admitted, transferred or discharged within four hours.
This is where gaming behaviour might be observed between the EDs and the EMS.
An assumption of this work is that some managerial decision making is involved
in choosing when to start blocking ambulances.
This is similar to~\cite{deo2011centralized}.

% Major contributions
The major contributions of this paper are:
\begin{itemize}
    \item A queueing model with 2 consecutive waiting spaces where one would 
    serve as a parking space for the ambulances
    \item Analytic performance measure formulas for the queueing model
    \item A 3-player game theoretic model between the EMS and two EDs
    \item Numerical experiments showing emergent behaviour of gaming between
    EDs and the EMS.
\end{itemize}
Specifically, our focus is on the construction of a 3-player game theoretic 
model between two queueing systems and a service that distributes individuals
to them. 
The resultant model is then used to explore the emergent dynamics between 
the three players.
This study explores two new concepts: getting performance measures for a new
queueing theoretic model with a parking space and a service centre and
using a learning algorithm to model the emergence of behaviour.
The developed theoretical model is illustrated through the application to 
a healthcare system of two EDs and the EMS, exploring the inefficiencies that 
emerge and ways to apply some incentive mechanisms to improve them.
The EDs are modelled as two queueing systems each with a tandem buffer and a 
service centre.
The performance measures are then used as the utilities of the game.
The novelty of the queueing model here is a contribution not only the game 
theoretic literature but also to the queueing theoretic literature.
To the authors knowledge, no such model of a tandem queueing model with a pair 
of parameters for the buffer has been considered.
Although our research is motivated by the particular EMS-ED example, our 
developed modelling framework and behavioural insights has application to 
similar systems across a range of sectors and settings.

>>>>

There are numerous news articles that focus on the complexity that arises when
ambulances stay blocked outside of the hospital for a long amount of 
time~\cite{dailyrecords, staffordshirelive}.
Some news reports comment on the long idle time of ambulances when
they are not in use~\cite{herefordtimes} and there are several reports of 
examples where this became an issue for new patients~\cite{southwalesargus}
and paramedics~\cite{bbcwales}.

% ED + EMS application
This paper aims to describe the EMS-ED interface using a game theoretic model
informed by an underlying queueing model.
The model describes the situation where an ambulance service would
have to distribute its patients between two EDs.
The two EDs can be thought of as two queueing systems and the EMS as a 
distributor that distributes patients to them, aiming to minimise some
performance measure.
The patients that are distributed by the EMS arrive at the hospital via an 
ambulance and are then either offloaded at the ED or stay blocked outside in 
the ambulance.
Whether or not the ambulance and its patient stay blocked is determined by 
the threshold that the given ED chooses to play.
A high threshold indicates that the ED accepts ambulance patients even if it is 
relatively full, while a low threshold means that the ED blocks ambulances more 
frequently.

In the United Kingdom, the National Heath Services (NHS) sets some regulations 
on ED performance.
One of these regulations is that 95\% of patients that arrive at the ED should 
be admitted, transferred or discharged within four hours.
This is where gaming behaviour might be observed between the EDs and the EMS.
An assumption of this work is that some managerial decision making is involved
in choosing when to start blocking ambulances.
This is similar to~\cite{deo2011centralized}.

% Major contributions
The major contributions of this paper are:
\begin{itemize}
    \item A queueing model with two consecutive waiting spaces where one would 
    serve as a parking space for the ambulances.
    \item Analytic performance measure formulas for the queueing model.
    \item A 3-player game theoretic model between the EMS and two EDs.
    \item Numerical experiments showing emergent behaviour of gaming between
    EDs and the EMS.
\end{itemize}
Specifically, our focus is on the construction of a 3-player game theoretic 
model between two queueing systems and a service that distributes individuals
to them. 
The resultant model is then used to explore the emergent dynamics between 
the three players.
This study explores two new concepts: getting performance measures for a new
queueing theoretic model with a parking space and a service centre, and
using a learning algorithm to model the emergence of behaviour.
The developed theoretical model is illustrated through the application to 
a healthcare system of two EDs and the EMS, exploring the inefficiencies that 
emerge and ways to apply some incentive mechanisms to improve them.
The EDs are modelled as two queueing systems each with a tandem buffer and a 
service centre.
The performance measures are then used as the utilities of the game.
The novelty of the queueing model here is a contribution not only the game 
theoretic literature but also to the queueing theoretic literature.
To the authors knowledge, no such model of a tandem queueing model with a pair 
of parameters for the buffer has been previously considered.

### 2022-04-05 sections/model_overview/methodology/main.tex
Changes after second supervisory review (#45)
* Add changes to response to reviewers

* Fix square in general markov model image

* Add another keyword

* Apply changes to intro and lit review

* Apply changes to queueing mode section

* Apply changes to game theoretic mode section

* Apply changes to conclusion

* Minor changes to appendix

* Compile PDF

The utilities to each player can be represented by 3 matrices; the two payoff 
matrices of the normal form game and the routing matrix.
The payoff matrices and their utilities are defined by equations 
(\ref{eq:payoff_entry}) and (\ref{eq:payoff_matrices}).

The routing matrix holds the values \((p_A, p_B)\) which are the proportion 
of ambulance patients to send to queueing systems \(A\) and \(B\).
Each pair \((p_A, p_B)\) can be calculated using equation 
(\ref{eq:obj_distributor}) and is essentially
a best response to the actions of the hospitals.
Thus, using equation (\ref{eq:obj_distributor}) for all possible sets of 
thresholds we can get the full routing matrix that consists of the proportions
to send to hospital A (\(p_A\)) and to hospital B (\(p_B\)).

@@

Consider Figure \ref{fig:imperfect_info_game} and the flow of the game that was
described (i.e. \(H_A, H_B \rightarrow D\)).
Due to the fact that the payoff matrices \(A\) and \(B\) depend on the routing 
matrix \(R\) the entries of the matrices are calculated in a backwards way 
(\(D \rightarrow H_A, H_B\)). 
This is done using backwards induction. 
For each action choice of the hospitals, first solve the game from the
ambulance's point of view.
This in effect results in a 2-player normal form game representing the
hospital's point of view.
Thus, for every pair of strategies \(T_A, T_B\), the values of \(p_A\) and 
\(p_B\) that satisfy equation (\ref{eq:obj_distributor}) are found
numerically using Brent's bisection algorithm~\cite{brent1973algorithms}.
Each pair \((p_A, p_B)\) corresponds to the best response of the ambulance
service to the two hospitals' played strategies.
Finally, using the routing matrix, equation (\ref{eq:payoff_entry}) can also be
used to populate the payoff matrices of the hospitals since we now know the 
arrival rate of each hospital.

>>>>

The utilities to each player can be represented by 3 matrices: the two payoff 
matrices of the normal form game and the routing matrix.
The payoff matrices and their utilities are defined by equations 
(\ref{eq:payoff_entry}) and (\ref{eq:payoff_matrices}).

The routing matrix holds the values \((p_A, p_B)\) which are the proportion 
of ambulance patients to send to queueing systems \(A\) and \(B\).
Each pair \((p_A, p_B)\) can be calculated using equation 
(\ref{eq:obj_distributor}) and is essentially
a best response to the actions of the hospitals.
Thus, using equation (\ref{eq:obj_distributor}) for all possible sets of 
thresholds, we can get the full routing matrix \(R\) (equation
\ref{eq:routing_matrix})
that consists of the proportions
to send to hospital A (\(p_A\)) and to hospital B (\(p_B\)).

@@

Now consider Figure \ref{fig:imperfect_info_game} and the flow of the game that
was described (i.e. \(H_A, H_B \rightarrow D\)).
Due to the fact that the payoff matrices \(A\) and \(B\) depend on the routing 
matrix \(R\) the entries of the matrices are calculated in a backwards way 
(\(D \rightarrow H_A, H_B\)). 
This is done using backwards induction. 
For each action choice of the hospitals, first solve the game from the
ambulance's point of view.
This in effect results in a 2-player normal form game representing the
hospital's point of view.
Thus, for every pair of strategies \(T_A, T_B\), the values of \(p_A\) and 
\(p_B\) that satisfy equation (\ref{eq:obj_distributor}) are found
numerically using Brent's bisection algorithm~\cite{brent1973algorithms}.
Each pair \((p_A, p_B)\) corresponds to the best response of the ambulance
service to the two hospitals' played strategies.
Finally, using the routing matrix, equation (\ref{eq:payoff_entry}) can also be
used to populate the payoff matrices of the hospitals since we now know the 
arrival rate of each hospital.

### 2022-04-05 sections/model_overview/numerical_results/main.tex
Changes after second supervisory review (#45)
* Add changes to response to reviewers

* Fix square in general markov model image

* Add another keyword

* Apply changes to intro and lit review

* Apply changes to queueing mode section

* Apply changes to game theoretic mode section

* Apply changes to conclusion

* Minor changes to appendix

* Compile PDF

In order to analyse how efficient the strategies played at every iteration are, 
the concept of the price of anarchy is used.
Price of anarchy is a measure that is used to quantify the efficiency of a
behaviour \cite{roughgarden2005selfish}.
In other words the price of anarchy is the worst-case equilibria measure and it 
is defined as:

@@

The learning algorithm reaches a stable pair of 
strategies where \(T_A = 5\) and \(T_B = 6\). After a number of iterations the
price of anarchy for both players stabilises and barely increases until the 
final iteration.

\textbf{Increasing \(\lambda_2\):}
Figure \ref{fig:ard_lambda_2} shows a similar run of the
algorithm but when the strategies begin to stabilise an increase in the
arrival rate of ambulances occurs (i.e \( \lambda_2 = 24 \)).

@@

By increasing \(\lambda_2\) there is no change as to how players behave
(\(T_A = 5, T_B = 6\)), but the efficiency of the system does change. 
There is a decline in the price of anarchy of the blocking time which at first 
glance indicates that upon flooding the system it becomes the loss in efficiency
due to rational individual behaviour decreases. 
This is non-sensical though.
What it really shows is that the steep increase in \( \lambda_2 \) leaves 
the system unable to cope regardless of the decisions made.

@@

\textbf{Incentivising players:}
From Figures \ref{fig:ard_lambda_2} and \ref{fig:ard_num_of_servers} it can be
seen that we can change some parameters of the model to make it more efficient.
The approach used on Figure \ref{fig:ard_penalty} is slightly different than 
the previous cases.
Once the played strategies in asymmetric replicator dynamics strategies 
converge the payoff matrices of the two are scaled in such a way so that the 
utilities of the selected strategy are penalised. This corresponds to a precise
policy change where more societally beneficial behaviours are incentivised.

>>>>

In order to analyse how efficient the strategies played at every iteration are, 
the concept of the price of anarchy is used.
Price of anarchy (PoA) is a measure that is used to quantify the efficiency of
a behaviour \cite{roughgarden2005selfish}.
In other words the price of anarchy is the worst-case equilibria measure and it 
is defined as:

@@

The learning algorithm reaches a stable pair of 
strategies where \(T_A = 5\) and \(T_B = 6\). 
Thereafter, the
price of anarchy for both players stabilises and barely increases.

\textbf{Increasing \(\lambda_2\):}
Figure \ref{fig:ard_lambda_2} shows a similar run of the
algorithm but when the strategies begin to stabilise, an increase in the
arrival rate of ambulances occurs (i.e. \( \lambda_2 = 24 \)).

@@

By increasing \(\lambda_2\) there is no change as to how players behave
(\(T_A = 5, T_B = 6\)), but the efficiency of the system does change. 
There is a decline in the price of anarchy of the blocking time which at first 
glance indicates that upon flooding the system it becomes the loss in efficiency
due to rational individual behaviour decreases. 
This is non-sensical however.
What it really shows is that the steep increase in \( \lambda_2 \) leaves 
the system unable to cope regardless of the decisions made.

@@

\textbf{Incentivising players:}
From Figures \ref{fig:ard_lambda_2} and \ref{fig:ard_num_of_servers} it is
observed that we can change some parameters of the model to make it more
efficient.
The approach used on Figure \ref{fig:ard_penalty} is slightly different than 
the previous cases.
Once the played strategies in asymmetric replicator dynamics strategies 
converge, the payoff matrices of the two are scaled in such a way so that the 
utilities of the selected strategy are penalised. This corresponds to a precise
policy change where more societally beneficial behaviours are incentivised.
